---
title: "Visualization Part 2"
output: html_notebook
---

## A working example

The previous notebook was a simple introduction to the syntax for constructing a `ggplot` object.  This notebook will introduce a more meaningful example using landcover data from the NSF National Ecological Observation Network (NEON) project.  For more information you can look at the [NEON website](http://www.neonscience.org)

Load the tidyverse library

```{r libraries}
library(tidyverse)
```


For this project we are going to use vegetation structure data from the San Juaquin Experimental Range in California.

### Load the data

```{r load-structure-data}
sjer_veg <- read_csv("data/D17_2013_SJER_vegStr.csv")
# View(sjer_veg)
sjer_veg
```

The loaded data contain observations for vegetation structure for individual plots at the site.  We can use our data wrangling skills to query the data, summarzing various fields of interest by plot.

```{r summary-queries}
sjer_veg %>% 
  group_by(plotid) %>% 
  summarise(mean(maxcanopydiam))
```

We can quickly visualize this in a single piped operation.

```{r summary-viz}
sjer_veg %>% 
  group_by(plotid) %>% 
  summarise(m = mean(maxcanopydiam)) %>% 
  ggplot(aes(x = plotid, y = m)) +
    # geom_bar(stat = "identity") + # a column plot for each entry
    geom_col() + # geom_col is a convenience function for using values instead of counts
    ggtitle("SJER Distribution of Canopy Diameter by Plot") +
    xlab("Plot ID") + 
    ylab("Max Diameter (m)") +
    theme_bw() + #change the theme
    theme(plot.margin=unit(c(.5,1,.5,1), "cm"), #adjust the margin (top, right, bottom, left)
          axis.text.x = element_text(angle = -60, hjust = -0.1)) # rotate the axis labels

```


The benefit of the tidyverse is how well the libraries work together to streamline data prep, query, and visualizations.  Let's try another example.  The data folder contains a file with a description of the fields in the vegStructure data.  There is a field called `livingcanopy` describing the estimated percent of the canopy that is living.  We can use that as a color gradient for a plot of canopy diameter compared to stem height.

```{r height-diameter}

sjer_veg %>% 
  select(maxcanopydiam, stemheight, livingcanopy) %>% 
  ggplot(aes(x = stemheight, y = maxcanopydiam)) + 
    geom_point(aes(color = livingcanopy), size = 3) + # change the point size
    scale_color_gradient(low = "red", high = "green", 
                         name = "Percent Living", # add a label for the legend
                         labels = scales::unit_format("%"), # reformat the units
                         breaks = c(0, 50, 100)) + # change the breakpoints in the scale
    geom_abline( size = 1, color = "blue") + # add a regression line
    ggtitle("SJER Relationship of Percent Living Canopy by Canopy Structure") + # add title
    xlab("Stem Height (m)") + # add x label
    ylab("Max Canopy Diameter (m)") + # add y label
    theme_minimal() #change the theme
    

# how can we improve this plot?
```



