---
title: "Visualizations Assignment 2"
author: "Chase LaDue"
date: "4/9/2018"
output: html_document
---

The goal of this assignment is to generate a visualization using my own data. For this, I will be using elephant chemosensory data described in a previous assignment (located in EVPP505-projects/data_wrangle/ele_data_wrangle.Rmd).

```{r libraries}
library(tidyverse)
library(RColorBrewer)
```

```{r import-data}
ElePheromoneData <- read_csv("data/ElePheromoneData.csv")
```

```{r custom-theme}
custom.theme <- theme(panel.grid.major = element_blank(), # these lines gets rid of distracting grid in background
                      panel.grid.minor = element_blank(),
                      panel.background = element_rect(fill="white"),
                      axis.line = element_line(colour="black"), # this changes axes lines to black
                      axis.text.x = element_text(colour="black", size=10), # changes axes text (numbers)
                      axis.text.y = element_text(colour="black", size=10),
                      axis.title.x = element_text(colour="black", size=12, face="bold"), # changes axes labels
                      axis.title.y = element_text(colour="black", size=12, face="bold"),
                      plot.title = element_text(face="bold",hjust=0.5), # centers plot title
                      legend.title = element_text(face="bold", size=10), # changes legend title text
                      legend.title.align = 0.5, # centers legend title
                      legend.text = element_text(size=8), # changes legend text
                      legend.background = element_rect(fill="#F6F6F6", colour = "black")) # adds background and border to legend
```


As a reminder, I've already made a graph of the average age for each sex-experience group, but this was using Base R. This is what it looked like:

```{r base-r-graph}

# data wrangling
age.summary.data <- ElePheromoneData %>% 
  group_by(Experience) %>% 
  summarize(Number.Eles = (n()/6), Average.Age = mean(Age), Median.Age = median(Age))

# graph labels
sex.exp.group <- c("Female, Exp.", "Female, Naive", "Male, Exp.", "Male, Naive")

# create plot
barplot(age.summary.data$Average.Age, 
        names.arg = sex.exp.group, 
        col = c("lightpink", "lightpink", "lightskyblue", "lightskyblue"), 
        ylab = "Average Age (yr)", 
        xlab = "Sex-Experience Group", 
        main = "Age Summary Data",
        ylim = c(0,40)
        )
```

However, this graph may be an oversimplification of these data. I'll use `ggplot2` to better interpret the data. First, I need to separate the "Experience" column in `ElePheromoneData` into two separate columns: one for sex and one for breeding history

```{r data-separate}

ElePheromoneData <- ElePheromoneData %>% 
  mutate(sex = case_when(
    Experience %in% c("F-Naive", "F-Experienced") ~ "Female",
    Experience %in% c("M-Naive", "M-Experienced") ~ "Male",
    TRUE ~ "Unknown")
    ) %>% 
  mutate (breeding.history = case_when(
   Experience %in% c("F-Naive", "M-Naive") ~ "Naive",
   Experience %in% c("F-Experienced", "M-Experienced") ~ "Experienced",
   TRUE ~ "Unknown")
   )

# Make sure everything worked correctly:
# View(ElePheromoneData)

```

```{r base-plot}
ggplot(ElePheromoneData, aes(x = breeding.history, y = c.Positive, fill = Treatment)) +
  geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
  #geom_errorbar(aes(ymin=c.Positive-se, ymax=c.Positive+se), position = "dodge") +
  facet_wrap(~ sex + Chemical)
```

This would be more informative with error bars, but I'll need to calculate the values for these.

```{r calculate-sterr}

graph.data <- ElePheromoneData %>% 
  group_by(Chemical, sex, breeding.history, Treatment) %>% 
  summarize(Average.Behav = mean(c.Positive), StErr = (sd(c.Positive)/sqrt(n())))

```

Now that I've created a new tibble to calculate standard error, I need to go back to the base plot I created above and substitute the correct values. Then, I can add error bars.

```{r add-error-bars}
a.1 <- ggplot(graph.data, aes(x = breeding.history, y = Average.Behav, fill = Treatment)) +
  geom_bar(position = "dodge", stat = "summary", fun.y = "mean") +
  geom_errorbar(aes(ymin=Average.Behav-StErr, ymax=Average.Behav+StErr), position = position_dodge(0.9), width = 0.2) +
  facet_wrap(~ sex + Chemical) + custom.theme

# a.1
```

These plots make it easy to interpret the data. Let's change the appearance to make it even easier

```{r labels}
a.2 <- a.1 + 
  xlab("Breeding history") +
  ylab("Proportion of time: positive behavior") +
  ggtitle("Effect of pheromone treatment on positive behaviors") +
  scale_fill_discrete(labels=c("Pre-treatment", "Treatment", "Post-treatment"), name = "Experimental phase")
  
# a.2
```

Let's get rid of the junk around the facets so that they're easier to read, and don't interfere with the plots themselves.

```{r change-facets}

custom.theme <- custom.theme + theme(
  strip.text = element_text(face="bold"),
  strip.background = element_rect(fill="white")
)

a.3 <- a.2 + custom.theme

# a.3

```

Let's change to a color palette that's easier to read. Note: for the following code, need package `RColorBrewer` that was installed at beginning of this file.

```{r change-color-palette}

a.4 <- a.3 + scale_fill_brewer(palette="Pastel2", labels=c("Pre-treatment", "Treatment", "Post-treatment"), name = "Experimental phase")

a.4

```

I don't like that the bars aren't flush with 0 on the x-axis. We can manually define limits, and use `expand`, to fix this.

```{r reposition-bars}
a.5 <- a.4 + scale_y_continuous(limits = c(0,0.35), expand = c(0,0))

# a.5
```


This graph is easier to interpret now with all the changes I've made.
```{r final-graph}
final.graph <- a.5

final.graph
```

