---
title: "EVPP 505 R Class Project, Spring 2018"
Author: Stephanie Gordon 
output: html_notebook
---

This is an rnotebook to document class work for Spring 2018 EVPP 505 R class

Goals of project: Determine annual chemical output per facility, and find output as percent of total facilty flow

Steps: 

1. read in DMR data - check

2. join DMR data together - check

2. find number of unique faacilities - check

3. select important columns - check

4. sum total chem flow by facility - check 

5. find total chem as percent of annual actual flow 

```{r set-environments-and-libraries}
library(readr)
library(dplyr)
library(tidyr) #stringr is part of tidyverse
library(ggplot2)
library(plotrix)
```


```{r read-in-data}
DMR_1 <- read.csv("Data/DMR_2016_1.csv")
DMR_2 <- read.csv("Data/DMR_2016_2.csv", skip = 1)
DMR_3 <- read.csv("Data/DMR_2016_3.csv", skip = 1)
DMR_4 <- read.csv("Data/DMR_2016_4.csv", skip = 1)
DMR_5 <- read.csv("Data/DMR_2016_5.csv", skip = 1)
```


```{r stack-em-up}
DMR_All <- rbind(DMR_1, DMR_2, DMR_3, DMR_4, DMR_5)
head(DMR_All)
#slice = dplyr version
```

```{r unique-facilities}
#how many facilities in total? use later to confirm that the summaries are for all facilities
Facilities <- unique(as.data.frame(DMR_All$Facility.Name))
#count(Facilities)

#OR use this - doesn't save facility names, just displays number of unique records in a selected column - just a vector of the answer
n_distinct(DMR_All$Facility.Name)

#same result as above, but creates a tibble/dataframe that you can add columns etc. on to 
DMR_All %>% 
  summarise(n_distinct(Facility.Name))

```

```{r calculate-pollutant-load-per-facility}
DMR_cleaned <- DMR_All %>% 
  select(Facility.Name, Parameter.Description, Pollutant.Load..kg.yr., Max.Allowable.Load..kg.yr., Wastewater.Flow..MGal.yr.) %>%
  mutate(Wastewater.Flow.kg..yr. = 3.7854118*Wastewater.Flow..MGal.yr.) %>% #add column with annual flow in kg insteaed of gallons
  group_by(Facility.Name) %>% 
  summarize(PollutantLoadSum = sum(Pollutant.Load..kg.yr.)) %>% 
  filter(!is.na(PollutantLoadSum))

DMR_cleaned2 <- DMR_All %>% 
  select(Facility.Name, Parameter.Description, Pollutant.Load..kg.yr., Max.Allowable.Load..kg.yr., Wastewater.Flow..MGal.yr.) %>%
  mutate(Wastewater.Flow.kg..yr. = 3.7854118*Wastewater.Flow..MGal.yr.) %>% 
  group_by(Facility.Name) %>% 
  summarize(wastewaterflowMkg = mean(Wastewater.Flow.kg..yr.))

#merge to create single datafram with jsut the pollutant load and flow variables
DMR_cleaned_merge <- merge(x = DMR_cleaned, y = DMR_cleaned2, by.x = "Facility.Name", by.y = "Facility.Name", all.x = TRUE)

#remove ND flow records
DMR_cleaned_merge_noNA <- DMR_cleaned_merge %>%
  filter(!is.na(wastewaterflowMkg))

#double check a single facility to confirm results from DMR_cleaned (first part above)
filter(DMR_All, Facility.Name == '#1 STATE UPRR, HORNE BROTHERS') %>% 
  select(Facility.Name, Pollutant.Load..kg.yr.) %>% 
  summarise(sum(Pollutant.Load..kg.yr.))
```

```{r plots-histograms-and-barplots}
#try to plot pollutant loads
DMR_hist <- ggplot(data = DMR_cleaned_merge, aes(DMR_cleaned_merge$PollutantLoadSum)) +
  geom_histogram() +
  stat_bin(bins = 40) 
DMR_hist

gap.barplot(DMR_cleaned_merge$PollutantLoadSum, gap=c(200,2000),horiz=F)
gap.barplot(DMR_cleaned_merge$PollutantLoadSum, gap=c(200,4000),horiz=T)


#switch to flow
gap.plot(DMR_cleaned_merge_noNA$wastewaterflowMkg, 
         gap.axis = "y", 
         gap = c(200,3800),
         xlab = "Wastewater Flow (Mkg)",
         ylab = "Count")

hist(DMR_cleaned_merge_noNA$wastewaterflowMkg,
        xlab = "Wastewater Flow (Mk/yr)")
  
min(DMR_cleaned_merge_noNA$wastewaterflowMkg)
max(DMR_cleaned_merge_noNA$wastewaterflowMkg)
mean(DMR_cleaned_merge_noNA$wastewaterflowMkg) #use to set igh/low values for split histogram
median(DMR_cleaned_merge_noNA$wastewaterflowMkg)

#facet to wrap histogram in 2 plots to show smaller values
#https://stackoverflow.com/questions/10504804/put-a-break-in-the-y-axis-of-a-histogram?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa


DMR_cleaned_merge_noNA$lowhigh <- ifelse(DMR_cleaned_merge_noNA$wastewaterflowMkg >= 5457.211,  c("High"), c("Low")) 
DMR_cleaned_merge_noNA %>%
  ggplot(aes(x = wastewaterflowMkg)) + 
  geom_histogram(colour = "orange", fill = "orange") + 
  facet_wrap(~ lowhigh, scales = "free") +
  labs(title = "Flow Histogram with Break in Y-axis") +
  theme_classic()

DMR_cleaned_merge_noNA %>%
  ggplot(aes(x = wastewaterflowMkg)) + 
  geom_histogram(colour = "blue", fill = "blue") + 
  facet_wrap(~ lowhigh) +
  labs(title = "Flow Histogram with Break in Y-axis") +
  theme_classic()
 
 
length(which(DMR_cleaned_merge_noNA$lowhigh == "Low"))  #2690
```


```{r examples-1}
# gap plots : https://www.rdocumentation.org/packages/plotrix/versions/3.7/topics/gap.plot 
 twogrp<-c(rnorm(5)+4,rnorm(5)+20,rnorm(5)+5,rnorm(5)+22)
 gpcol<-c(2,2,2,2,2,3,3,3,3,3,4,4,4,4,4,5,5,5,5,5)
 
 gap.plot(twogrp,gap=c(8,16),xlab="Index",ylab="Group values",
  main="Gap on Y axis",col=gpcol)
 
 gap.plot(twogrp,rnorm(20),gap=c(8,16),gap.axis="x",xlab="X values",
  xtics=c(4,7,17,20),ylab="Y values",main="Gap on X axis with added lines")
 
 gap.plot(c(seq(3.5,7.5,by=0.5),seq(16.5,22.5,by=0.5)),
  rnorm(22),gap=c(8,16),gap.axis="x",type="l",add=TRUE,col=2,)
 
 gap.plot(twogrp,gap=c(8,16,25,35),
 xlab="X values",ylab="Y values",xlim=c(1,30),ylim=c(0,42),
 main="Test two gap plot with the lot",xtics=seq(0,30,by=5),
 ytics=c(4,6,18,20,22,38,40,42),
 lty=c(rep(1,10),rep(2,10)),
 pch=c(rep(2,10),rep(3,10)),
 col=c(rep(2,10),rep(3,10)),
 type="b")
 
 gap.plot(21:30,rnorm(10)+40,gap=c(8,16,25,35),add=TRUE,
  lty=rep(3,10),col=rep(4,10),type="l")
```

```{r example-2}
twogrp<-c(rnorm(10)+4,rnorm(10)+20)
 gap.barplot(twogrp,gap=c(8,16),xlab="Index",ytics=c(3,6,17,20),
  ylab="Group values",main="Barplot with gap")
 gap.barplot(twogrp,gap=c(8,16),xlab="Index",ytics=c(3,6,17,20),
  ylab="Group values",horiz=TRUE,main="Horizontal barplot with gap")
```


```{r}
ggplot(DMR_cleaned_merge_noNA, aes(x = PollutantLoadSum, y = wastewaterflowMkg)) +
  geom_point()

DMR_cleaned_merge_noNA_All <- DMR_cleaned_merge_noNA %>%
  mutate(pollutantpercentofflow = ((PollutantLoadSum/wastewaterflowMkg)*100)) %>%
  merge(DMR_All, by.x = "Facility.Name", by.y = "Facility.Name") %>%
  ggplot(aes(x = PollutantLoadSum, y = wastewaterflowMkg, color = State)) +
  geom_point(stat = "unique") +
  xlab("Total Pollutant Load (kg/yr)") + 
  ylab("Wasterwater Flow (Mkg/yr)")
DMR_cleaned_merge_noNA_All
```




